Discrete Ordinates

### 2017-01-19 docprep/draft1.tex
un/collided flux clarification
Added a sentence about the indexing scheme
Clarified the Legendre part and omitted a bunch of old stuff

\documentclass{article}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[toc,page]{appendix}
\usepackage{listings}
\usepackage{color}
\usepackage{listings}
\usepackage{fullpage}

\numberwithin{equation}{subsection}

@@

@@

which becomes

@@

Legendre polynomials are solutions to
\begin{equation}
P_n(x) = \frac{1}{2 \pi i} \oint (1-2\xi x + \xi^2)^{-1/2} \xi^{-n-1} d\xi
\end{equation}

\begin{equation}
P_n(x) = \frac{1}{2^n n!}\frac{d^n}{dx^n}(x^2-1)^n
\end{equation}

@@

@@

@@

Both Legendre polynomials and Associated Legendre Polynomials can be calculated using the Boost C++ library.

@@

First, we assume that Laplace's equation is of the seperable form $f(\rho, \theta, \phi) = R(\rho)Y(\theta, \phi)$. Equation \ref{eq_laplace_spherical} can then be rewritten as

\begin{equation} \label{eq_spherical1}
\begin{split}
\nabla^2 f(\rho, \theta, \phi) = 
\frac{1}{\rho^2}
\frac{\partial}{\partial \rho} 
\left( \rho^2 \frac{\partial (R(\rho)Y(\theta, \phi))}{\partial \rho}\right) 
+ \\
\frac{1}{\rho^2 \sin \theta}
\frac{\partial}{\partial \theta}
\left( \sin \theta \frac{\partial (R(\rho)Y(\theta, \phi))}{\partial \theta}\right) 
+ 
\frac{1}{\rho^2 \sin^2 \theta} 
\frac{\partial^2 (R(\rho)Y(\theta, \phi))}{\partial \phi^2} = 0
\end{split}
\end{equation}

\noindent
which simplifies to
\begin{equation} \label{eq_spherical2}
\begin{split}
\frac{Y(\theta, \phi)}{\rho^2}
\frac{\partial}{\partial \rho} 
\left( \rho^2 \frac{\partial R(\rho)}{\partial \rho}\right) 
+ 
\frac{R(\rho)}{\rho^2 \sin \theta}
\frac{\partial}{\partial \theta}
\left( \sin \theta \frac{\partial Y(\theta, \phi)}{\partial \theta}\right) 
+ \\
\frac{R(\rho)}{\rho^2 \sin^2 \theta} 
\frac{\partial^2 Y(\theta, \phi)}{\partial \phi^2}
\end{split} = 0
\end{equation}

\noindent
Multiplying Eq. \ref{eq_spherical2} by $\rho^2 / (R(\rho)Y(\theta, \phi))$ yields

\begin{equation}\label{eq_spherical3}
\begin{split}
\frac{1}{R(\rho)}
\frac{\partial}{\partial \rho} 
\left( \rho^2 \frac{\partial R(\rho)}{\partial \rho}\right) 
+ 
\frac{1}{Y(\theta, \phi) \sin \theta}
\frac{\partial}{\partial \theta}
\left( \sin \theta \frac{\partial Y(\theta, \phi)}{\partial \theta}\right) 
+ \\
\frac{1}{Y(\theta, \phi) \sin^2 \theta} 
\frac{\partial^2 Y(\theta, \phi)}{\partial \phi^2} = 0
\end{split}
\end{equation}

Equation \ref{eq_spherical3} can be split into two parts, a function of $\rho$ alone and a component that is a function of $\theta$ and $\phi$ alone. The component that is a function of $\theta$ and $\phi$ can be moved to the other side of the equation. Since two functions of different variables are equal, they must each be equal to some constant, $\lambda$. Therefore, Eq. \ref{eq_spherical3} can be rewritten as two equations:

\begin{equation}\label{eq_spherical_r}
\frac{1}{R(\rho)}
\frac{\partial}{\partial \rho} 
\left( \rho^2 \frac{\partial R(\rho)}{\partial \rho}\right) = \lambda
\end{equation}

\begin{equation}\label{eq_spherical_tp}
\frac{1}{Y(\theta, \phi) \sin \theta}
\frac{\partial}{\partial \theta}
\left( \sin \theta \frac{\partial Y(\theta, \phi)}{\partial \theta}\right) 
+ \\
\frac{1}{Y(\theta, \phi) \sin^2 \theta} 
\frac{\partial^2 Y(\theta, \phi)}{\partial \phi^2} = -\lambda
\end{equation}

$Y(\theta, \phi)$ can be further seperated into $Y(\theta, \phi) = \Theta(\theta)\Phi(\phi)$ which produces

\begin{equation}\label{eq_spherical_y}
\begin{split}
\frac{1}{\Theta(\theta)\Phi(\phi) \sin \theta}
\frac{\partial}{\partial \theta}
\left( \sin \theta \frac{\partial (\Theta(\theta)\Phi(\phi))}{\partial \theta}\right) 
+ \\
\frac{1}{\Theta(\theta)\Phi(\phi) \sin^2 \theta} 
\frac{\partial^2 (\Theta(\theta)\Phi(\phi))}{\partial \phi^2} = -\lambda
\end{split}
\end{equation}

\noindent
which simplifies to

\begin{equation}\label{eq_spherical_y2}
\begin{split}
\frac{1}{\Theta(\theta) \sin \theta}
\frac{\partial}{\partial \theta}
\left( \sin \theta \frac{\partial \Theta(\theta)}{\partial \theta}\right) 
+ 
\frac{1}{\Phi(\phi) \sin^2 \theta} 
\frac{\partial^2 \Phi(\phi)}{\partial \phi^2} = -\lambda
\end{split}
\end{equation}

Multiplying Eq. \ref{eq_spherical_y2} by $\sin^2(\theta)$ and rearranging yields

\begin{equation}\label{eq_spherical_y3}
\begin{split}
\frac{\sin \theta}{\Theta(\theta)}
\frac{\partial}{\partial \theta}
\left( \sin \theta \frac{\partial \Theta(\theta)}{\partial \theta}\right) 
+ \lambda \sin \theta
 = \frac{-1}{\Phi(\phi)} 
\frac{\partial^2 \Phi(\phi)}{\partial \phi^2}
\end{split}
\end{equation}

As before, each side of Eq. \ref{eq_spherical_y3} is a function of a single variable, therefore, both sides are equal to some constant. In this case, we select $m^2$ to be  the constant variable. The produces

\begin{equation}\label{eq_spherical_t}
\begin{split}
\frac{\sin \theta}{\Theta(\theta)}
\frac{\partial}{\partial \theta}
\left( \sin \theta \frac{\partial \Theta(\theta)}{\partial \theta}\right) 
+ \lambda \sin \theta
 = m^2
\end{split}
\end{equation}

\begin{equation}\label{eq_spherical_p}
\begin{split}
\frac{1}{\Phi(\phi)} 
\frac{\partial^2 \Phi(\phi)}{\partial \phi^2} = -m^2
\end{split}
\end{equation}

Combining Eq. \ref{eq_spherical_r}, \ref{eq_spherical_t}, and \ref{eq_spherical_p} Laplace's equation in sperical coordinates can be expressed in a fully seperated form as

\begin{equation}\label{eq_spherical_sep}
\begin{split}
\frac{1}{R(\rho)}
\frac{\partial}{\partial \rho} 
\left( \rho^2 \frac{\partial R(\rho)}{\partial \rho}\right) 
&= \lambda \\
\frac{\sin \theta}{\Theta(\theta)}
\frac{\partial}{\partial \theta}
\left( \sin \theta \frac{\partial \Theta(\theta)}{\partial \theta}\right) 
+ \lambda \sin \theta
&= m^2 \\
\frac{1}{\Phi(\phi)} 
\frac{\partial^2 \Phi(\phi)}{\partial \phi^2}
&= -m^2
\end{split}
\end{equation}

It can be shown through a detailed anlysis that is outside the scope of this paper that $m$ and $\lambda$ must both be integers. $\lambda$ is further constrained in that the relation $\lambda = l(l+1)$ where $l \leq |m|$ and $l \in \mathbb{Z}$. Therefore, Eq. \ref{eq_spherical_sep} can be written as

\begin{equation}\label{eq_spherical_sep2}
\begin{split}
\frac{1}{R(\rho)}
\frac{\partial}{\partial \rho} 
\left( \rho^2 \frac{\partial R(\rho)}{\partial \rho}\right) 
&= l(l+1) \\
\frac{\sin \theta}{\Theta(\theta)}
\frac{\partial}{\partial \theta}
\left( \sin \theta \frac{\partial \Theta(\theta)}{\partial \theta}\right) 
+ l(l+1) \sin \theta
&= m^2 \\
\frac{1}{\Phi(\phi)} 
\frac{\partial^2 \Phi(\phi)}{\partial \phi^2}
&= -m^2 \\
& l, m \in \mathbb{Z} \\
& l \leq |m|
\end{split}
\end{equation}

% TODO
TODO - Discontinuous

It can be shown that the solution is stated

\begin{equation}\label{eq:spherical_harmonic}
Y_l^m(\theta, \varphi) = (-1)^{l}\sqrt{\frac{2l+1}{4 \pi}
\frac{(l-m)!}{(l+m)!}} P_l^m(\cos \theta) e^{im \varphi}
\end{equation}

@@

\begin{equation}\label{eq:spherical_harmonic_exp}
Y_l^m(\theta, \varphi) = (-1)^{l}\sqrt{\frac{2l+1}{4 \pi}
\frac{(l-m)!}{(l+m)!}} P_l^m(\cos \theta) \left[ \cos(m \varphi) + i \sin(m \varphi) \right]
\end{equation}

>>>>

\documentclass{article}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[toc,page]{appendix}
\usepackage{listings}
\usepackage{color}
\usepackage{listings}
\usepackage{fullpage}
\usepackage{mathtools}
\numberwithin{equation}{subsection}

@@

The indexing scheme is as follows: E, A, x, y, z, l, m

@@

which, for the uncollided flux, becomes

@@

Legendre polynomials are solutions to
\begin{equation}
P_n(x) = \frac{1}{2 \pi i} \oint \frac{1}{\sqrt{1-2\xi x + \xi^2}} \frac{1}{\xi^{n+1}} d\xi
\end{equation}

@@

Legencre polynomials are useful for approximating functions truncated at some point. Legendre polynomials are mutually orthogonal which gives them the property

\begin{equation}
f(x) = \sum_{n=0}^\infty P_n(x)
\end{equation}

where $f(x)$ is any arbitrary function. Often, like a Fourier series which uses sin and cos instead of Legendre functions, the series is cut off at some threshold as an approximation. As an example of this, consider the function $1.75x^3 + 0.6x^2+.21x-.06$ over the domain $[2,5]$ using the first four Legendre polynomials.

The coefficients are computed

\begin{align}
k_0 &= \int_2^5 P_0(x) f(x) dx = \int_2^5  f(x) dx\\
k_1 &= \int_2^5 P_1(x) f(x) dx = \int_2^5 x f(x) dx\\
k_2 &= \int_2^5 P_2(x) f(x) dx = \int_2^5 \frac{1}{2}(3x^2-1) f(x) dx\\
k_3 &= \int_2^5 P_3(x) f(x) dx = \int_2^5 \frac{1}{2}(5x^3 - 3x) f(x) dx\\
\end{align}

These values are then used to reconstruct the original function by

\begin{equation}
f(x) \approx k_0 P_0(x) + k_1 P_1(x) + k_2 P_2(x) + k_3 P_3(x)
\end{equation}

@@

\subsection{Associated Legendre Polynomials}\label{sec:assoc_leg}

The associated Legendre polynomials are solutions to the equation

@@

The associated Legendre polynomials can be computed recusively by

\begin{align}
P_m^m(x) &= (-1)^m(2m-1)!!(1-x^2)^(m/2) \\
P_{m+1}^m(x) &= x(2m+1)P_m^m(x) \\
(l-m) P_l^m(x) &= x(2l-1)P_{l-1}^m(x) - (l+m-1)P_{l-2}^m(x)
\end{align}

For example, $5!! = 5 \times 3 \times 1 = 15$ and $6!! = 6 \times 4 \times 2 = 48$.

@@

Combining the associated Legendre polynomials and the sin/cos series, the spherical harmonics equations can be derived. They are orthogonal in both directions.

\begin{equation}\label{eq:spherical_harmonic}
Y_l^m(\theta, \phi) = N_l^{|m|} P_l^{|m|}(\cos \theta) e^{i m \phi}
\end{equation}

@@

\begin{equation}\label{eq:spherical_harmonic_exp}
Y_l^m(\theta, \varphi) = N_l^{|m|} P_l^m(\cos \theta) \left[ \cos(m \phi) + i \sin(m \phi) \right]
\end{equation}

The normalization constant $N_l^m$ is defined as

\begin{equation}
N_l^m = \sqrt{\frac{2l+1}{4 \pi} \frac{(l-m)!}{(l+m)!}}
\end{equation}

In many cases, only the real part ($y_l^m$) of the complex spherical harmonics ($Y_l^m$) are necessary.

\begin{equation}
y_l^m =
\begin{cases*}
\sqrt{2} N_l^m \cos(m\phi) P_l^m(\cos(\theta)) \\
N_l^0 P_l^0(\cos(\theta)) \\
\sqrt{2} N_l^{|m|} \sin(|m| \phi) P_l^{|m|}(\cos(\theta))
\end{cases*}
\end{equation}
